---
title: "pgrid: Access PRIO GRID raster data in R"
author: "Philipp Hunziker"
date: "December 9, 2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# pgrid: Access PRIO GRID raster data in R

The pgrid package allows conveniently creating raster data sets (more specifically: [VeloxRaster objects](https://github.com/hunzikp/velox)) from the PRIO GRID raw data as distributed at <http://grid.prio.org>. 

## Installation

Install pgrid straight from github:

```{r eval = FALSE}
require(devtools)
devtools::install_github('hunzikp/pgrid')
```

## Usage

### Get and link the data

Before using the pgrid package you need to download the PRIO GRID data in csv format from the  [PRIO GRID download portal](http://grid.prio.org/#/download). 
You don't need to download all the available variables, but obviously, when using pgrid, you'll only have access to those variables you downloaded.

Once you've downloaded the static and/or yearly PRIO GRID data you're interested in, you need to let pgrid know where those files are located. To do so, use the `setDataPath` function, e.g. as follows:

```{r}
library(pgrid)

## Set paths
setDataPath(staticPath = '/home/hunzikp/Data/priogrid/PRIO-GRID Static Variables - 2017-12-08.csv',
            yearlyPath = '/home/hunzikp/Data/priogrid/PRIO-GRID Yearly Variables for 1946-2014 - 2017-12-08.csv')
```


### Get information about variables

To see what PRIO GRID variables the pgrid package has access to, use the `getVarNames` function. If the `details` option is turned on, the function returns additional information about variable sources, etc.

Here we inquire what *static* PRIO GRID variables pgrid has access to:
```{r}
## Get info about static variables
static.df <- getVarNames(type = 'static', details = TRUE)
names(static.df)
head(static.df[,c('name', 'displayName', 'unit')])
```
And equivalently, what *yearly* variables pgrid has access to:
```{r}
## Get info about yearly variables
yearly.df <- getVarNames(type = 'yearly', details = TRUE)
head(static.df[,c('name', 'displayName', 'unit')])
```

### Get raster data

The `getPrioRaster` function returns PRIO GRID data as rasters. 
In the following, we construct a raster containing distances to the nearest urban center in one band (a static variable), and nightlight emission data for all available years between 2005 and 2010 in the other bands.
Note that `getPrioRaster` returns a list with two elements: (1) the actual raster data, and (2) a data frame containing meta data. 

```{r}
## Get raster data
prio.ls <- getPrioRaster(names = c('ttime_mean', 'nlights_calib_mean'), years = 2005:2010)
prio.vx <- prio.ls$raster
meta.df <- prio.ls$meta
```

The meta data frame links the requested variables to the raster bands:

```{r}
## Show meta data
head(meta.df)
```

The raster data is returned as a [VeloxRaster object](https://github.com/hunzikp/velox). If we want to plot it, we can cast it as a `RasterStack`. In the following, we use the excellent [rasterVis](https://oscarperpinan.github.io/rastervis/) package for plotting.

```{r message = FALSE, fig.height = 8, fig.width = 8}
## Cast raster data as RasterStack
prio.stk <- prio.vx$as.RasterStack()

## Plot nightlights in 2005/2010
library(raster)
library(rasterVis)
nightPalette <- colorRampPalette(c("black", 
                                   rgb(218,165,32,maxColorValue=255), 
                                   rgb(255,215,0,maxColorValue=255), 
                                   "white"), bias = 0.5)
nightTheme <- rasterTheme(region = nightPalette(20))
levelplot(prio.stk[[meta.df$band[meta.df$name == 'nlights_calib_mean' & meta.df$year %in% c(2005,2010)]]], 
          maxpixels = 720*360, 
          par.settings = nightTheme)
```







